Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2020-05-06 article.tex
Method is mostly written down

One might expect that the hydrophobic epitopes presented,
for a same hydrophobicity, are as likely to stem from 
membrane proteins TMHs or from soluble proteins hydrophobic (non-TMH) regions.
For MHC-I, however, it is found that the 9-mers stemming
from TMHs are presented more often than expected by 
chance \cite{bianchi2017}
\richel{but also when compensating for hydrophobicity?}.
For MHC-II, it is unknown which percentage of binders 
is derived from TMHs for all 13-mers present in a proteome
that have the same hydrophobicity 
\richel{
  $\mathcal{H}_{3,1}$
}.

@@

We used the same analysis as \cite{bianchi2017},
except that instead of a human reference proteome.
Bianchi and colleagues obtained a distribution of 
percentages of MHC-I epitopes overlap with TMHs in Homo sapiens
for the different HLA haplotypes, with an average of 5.3\%.
We obtained a similar distribution of percentages of MHC-I epitopes that 
overlap with TMHs for the different HLA haplotypes, but then applied to
our viral and bacterial proteome.

@@

For MHC-II, we do the same analysis as MHC-I, except for
MHC-II haplotypes. We picked the MHC-II alleles
that occur with a frequency of at least 20\% in
the human population (\cite{greenbaum2011functional},
see \ref{table:mhc2_haplotypes} for coverage),
resulting in \richel{approx 8} HLAs.
The threshold of 20\% is arbitrarily, yet chosen to result in 5 and 10 HLA variants.  
As \cite{bianchi2017}, we define the 5\% of polypeptides 
with the lowest IC50 values as binders. We then counted the percentage
of binders that have one or more amino acids that are part of a TMH.

@@

Figure \ref{fig:1} shows the percentages of epitopes overlapping 
with TMHs for a human, COVID-19 and Mycobacterium proteome.
The KS test to determine if these percentages are sampled from
a same distribution resulted in a p-value of \richel{unknown},
which makes us reject/accept the hypothesis that these percentages
are sampled from the same distribution.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_1/fig_bbbq_1_watermarked.png}
  \caption{
    \% epitopes overlapping with transmembrane helix
    for a human, COVID-19 and Mycobacterium proteome.
    \richel{
      the data underlying this figure has been simulated
    }
  }
  \label{fig:1}
\end{figure}

@@

@@

	% Table \ref{tab:results} shows the location and binding strength for the
	% tuberculosis proteome.

	% \begin{figure}[ht]
	%   \includegraphics[width=\textwidth]{bbbq_3/figure_1.png}
	%   \label{fig:1}
	% \end{figure}

	% \begin{figure}[ht]
	%   \includegraphics[width=\textwidth]{figure_1_5.png}
	%   \label{fig:1_5}
	% \end{figure}

	% \begin{figure}[ht]
	%   \includegraphics[width=\textwidth]{figure_3.png}
	%   \label{fig:3}
	% \end{figure}

	% \input{table_imoimo.latex}
	% % has label tab:results
	% \richel{
	%   I'd enjoy 
	%   (1) a row with 'expected by chance', 
	%   (2) using percentages instead,
	%   (3) merging inside and outside
	% }

	% The inference model with the highest evidence in the
	% TMH-only alignment was [yet unknown] and [also yet unknown]
	% for the non-TMH alignment. Individual model weights are shown
	% in tables \ref{tab:evidences_tmh} 
	% and \ref{tab:evidences_non_tmh}.

	% The Bayesian inference resulted in [a distribution of mutation rates],
	% as shown in [absent figure].
	% The ESSes of the Bayesian parameter estimates was above 200, exact values
	% are shown in tables \ref{tab:esses_tmh} and \ref{tab:esses_non_tmh}.

@@

	% \frans{
	%   De TMH evolutie data van COVID is missachien interresanter. Ik zie het als volgt. 
	%   COVID is een mooi voorbeeld van een virus 
	%   Myco van een bacterie
	%   In de bacterie is duidelijk dat tenmisnte voor MHCII TMHs zijn overgrepresenteerd virus niet. 
	%   Dus verhaal kan zijn dat TMHs belangrijk voot bacteirele afweer niet zozeer voor virussen. 
	% 
	%   Hiervoor moeten we MHC-I doen voor myco
	%   Kijken wat evolutie data doet voor virus dan evt kijken wat dit doet voor myco en vergelijken. 
	% }
	% 
	% We compared the mutation rates between the TMH and non-TMH part of
	% multiple mycobacterium species. Where we expect no variation 
	% in mutation rate for every TMH amino acid,
	% \richel{
	%   we can test this, but unsure if that would make sense
	% }
	% we know that non-TMH part will have regions of different evolutionary
	% conservation: functional domains, especially in protein-protein
	% interactions will be strongly conserved, due to an even more constrained
	% set of peptides that enable a certain function.

	% \richel{
	%   Note that most bacteria are opportunistic pathogens.
	%   Note that most bacteria are generalists.
	%   Note that most bacteria have different cell membranes (and walls), that
	%   may have different functional constraints than a human cell membrane
	% }

>>>>

One might expect that the hydrophobic epitopes presented,
for a same hydrophobicity, are as likely to stem from 
membrane proteins TMHs or from soluble proteins hydrophobic (non-TMH) regions.
For MHC-I, however, it is found that the 9-mers stemming
from TMHs are presented more often than expected by 
chance \cite{bianchi2017}
\richel{but also when compensating for hydrophobicity?}.
For MHC-II, it is unknown which percentage of binders 
is derived from TMHs for all 13-mers present in a proteome
that have the same hydrophobicity.

@@

To determine the percentages of MHC-I epitopes overlap
with TMHs, we used the same 
analysis as \cite{bianchi2017},
except for using a viral and bacterial proteome (instead 
of a human proteome, as used in the original study).
Bianchi and colleagues obtained a distribution of 
percentages of MHC-I epitopes overlap with TMHs in Homo sapiens
for the different HLA haplotypes, with an average of 5.3\%.
We obtained a similar distribution of percentages of MHC-I epitopes that 
overlap with TMHs for the different HLA haplotypes, but then applied to
our viral and bacterial proteome.

@@

To determine the percentages of MHC-II epitopes overlap, 
we do the same analysis as MHC-I, except for
MHC-II haplotypes. We picked the MHC-II alleles
that occur with a frequency of at least 20\% in
the human population (\cite{greenbaum2011functional},
see \ref{table:mhc2_haplotypes} for coverage),
resulting in \richel{approx 8} HLAs.
The threshold of 20\% is arbitrarily, yet chosen to result in 5 and 10 HLA variants.  
As \cite{bianchi2017}, we define the 5\% of polypeptides 
with the lowest IC50 values as binders. We then counted the percentage
of binders that have one or more amino acids that are part of a TMH.

@@

Figure \ref{fig:1} shows the percentages of MHC-I epitopes overlapping 
with TMHs for our human, viral and bacterial proteome.

The KS test to determine if MHC-I has the same percentage 
of epitopes overlapping with viral TMHs, 
compared to human TMHs \richel{$\mathcal{H}_{1,V}$},
resulted in a p-value of \richel{unknown},
which makes us [reject/accept] the hypothesis that these percentages
are sampled from the same distribution.

The KS test to determine if MHC-I has the same percentage 
of epitopes overlapping with bacterial TMHs, 
compared to human TMHs \richel{$\mathcal{H}_{1,B}$},
resulted in a p-value of \richel{unknown},
which makes us [reject/accept] the hypothesis that these percentages
are sampled from the same distribution.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{bbbq_1/fig_bbbq_1_watermarked.png}
  \caption{
    Percentage of MHC-I epitopes overlapping with TMHs
    for a human, viral and bacterial proteome.
    \richel{
      the data underlying this figure has been simulated
    }
  }
  \label{fig:1}
\end{figure}

@@

Figure \ref{fig:2} shows the percentages of MHC-II epitopes overlapping 
with TMHs for our human, viral and bacterial proteome.

@@

@@

\richel{
  Note that most bacteria are opportunistic pathogens.
  Note that most bacteria are generalists.
  Note that most bacteria have different cell membranes (and walls), that
  may have different functional constraints than a human cell membrane
}
